---
output:
  pdf_document: default
  html_document: default
---
## List in R
R list is the object which contains elements of different types – like strings, numbers, vectors and another list inside it. R list can also contain a matrix or a function as its elements. The list is created using the list () function in R. 

## How to create a list containing strings, numbers, vectors and a logical value in R?
Create list containing string, numbers, vectors and logical values using list() function. 
```{r}
list_data <- list("kathmandu", "pokhara", c(1,2,3), TRUE, 22.4) 
print(list_data) 
```

## How to name the list element in R ?
The list can be created using list() function in R. Named list is also created with the same function by specifying the names of the elements to access them. 
```{r}
list_data <- list("a"="kathmandu", "b"="pokhara", "c"=c(1,2,3), "d"=TRUE, "e"=22.4) 
print(list_data) 
```
## How to assess list element in R?
```{r}
list_data <- list("a"="kathmandu", "b"="pokhara", "c"=c(1,2,3), "d"=TRUE, "e"=22.4) 
# if name assign to list element then $ is use
print(list_data$a)
# to assess the list element using location
print(list_data[1])
# access element of vectors
print(list_data[[3]][[3]])
```
## How to manipulate list element in R ?
List manipulation occurs by create, read, update and delete in list. 
```{r}
# create list data
list_data <- list("ktm","pkr",c(1,2,3), 34)
# read_list data
print(list_data)
```
```{r}
# add new element to list
list_data[4] <- 30.5
print(list_data)
```
```{r}
# delete last element
list_data[4] <- NULL
list_data
```
## How to converts list to vectors in R?
Convert list to vector using unlist().
```{r}
print(unlist(list_data))
```




